---
title: "Introduction and Essentials"
author: "Victoria Liebsch-Aljawahiri"
date: "2022-09-08"
output: 
  html_document:
    theme: cerulean
    highlight: kate
    code_folding: hide
    df_print: paged
    html_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Perparations 

Load required libraries and custom functions.

```{r, message = FALSE}
library(tidyverse)

# Functions
# Theme for plot
cool_theme <- function() {
  theme_minimal() +
  theme(
    text = element_text(family = "Bookman", color = "gray25"),
    plot.subtitle = element_text(size = 12),
    plot.caption = element_text(color = "gray30"),
    plot.background = element_rect(fill = "gray95"),
    plot.margin = unit(c(5, 10, 5, 10), units = "mm")
  )
}
```

## Question 1 

A data frame has the variables of a data set as columns and the observations as rows. Basically, a data frame is a combination of data sets that contain different data types. The code below shows an example of a data frame. I have used the ```head()``` function to shorten how much of the data frame is printed out. 

```{r}
head(iris)
```

## Question 2 
Create a data frame with 3 variables. The 3 variables should be cars, mpg, cost.
The variable cars should contain 3 inputs. (Truck, Car, SUV)
The variable mpg should contain 3 inputs (11, 30, 24)
The variable cost should contain 3 inputs (45000,25000,35000)
The data frame is printed below.

```{r}
cars <- c("Truck", "Car", "SUV")
mpg <- c(11, 30, 24)
cost <- c(45000, 25000, 35000)
vehicle_info <- data.frame(cars, mpg, cost)
vehicle_info
```

## Question 2a

Select row 1 in column 3, what was selected?   The first row of the ```cost``` variable.
Select rows 1 through 3, what was selected?   The entire data frame. 
Select the last column, what was selected?   All rows of the ```cost``` variable.

```{r}
vehicle_info[1,3]
vehicle_info[1:3,]
vehicle_info[,3]
```


### Questions 3-5 will use the built-in dataset ```mtcars```.

## Question 3

Display the first 3 rows of the data set and display the last 5 rows. Hint: ```head()``` and ```tail()``` are helpful
functions.

```{r}
mtcars[1:3,]
mtcars[28:32,]
```

## Question 4 

Display the column names of two categorical variables and two continuous variables.

```{r}
?mtcars
str(mtcars)
colnames(mtcars)[2]   # categorical
colnames(mtcars)[8]   # categorical
colnames(mtcars)[1]   # continuous
colnames(mtcars)[3]   # continuous
```

## Question 5 

Type and execute the code ```ggplot(mtcars, aes(x=disp, y=mpg))``` in R studio. Does the code display a chart? If so, explain the chart that
was displayed; if the chart is not displayed, indicate what is missing and how would you update the code
to show a scatterplot showing the relationship between ```disp``` and ```mpg```.

```{r}
ggplot(mtcars, aes(x=disp, y=mpg)) +
  cool_theme()
```

The code above displays a blank chart. To fix this you must add the ```geom_plot()``` function to the code. Below is the updated code showing the relationship between ```disp``` and ```mpg```.

```{r}
ggplot(data = mtcars) +
  geom_point(mapping = aes(x = disp, y = mpg)) +
  cool_theme()
```


### Questions 6&7 will use the dataset ``mpg``.

## Question 6 

Create a scatterplot showing the relationship between ```displ``` and ```hwy```. Color each point based on the
number of cylinders (```cyl```). Explain the scatterplot.

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl)) +
  cool_theme() 
```

This scatter plot shows the relationship between three variables (```hwy```, ```displ```, ```cyl```) while still using a 2-D x,y plane. 

## Question 7

Copy the code for the scatter plot in question 6 and update it to apply facet_wrap using the ```drv``` column.
What does the resulting charts indicate about rear wheel drive vehicles when compared to 4wd and front-
wheel drive? Hint: use the help function to get more information on ```facet_wrap```, e.g. ```?facet_wrap```.

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl)) +
  facet_wrap(vars(drv)) +
  cool_theme()
```

The charts reveal that rear wheel drive vehicles have the lowest variety of number of cylinders and tend to have more engine displacement (in liters), when compared to front and 4 wheel drive vehicles.

